---
title: "NES-LTER zooplankton transect abundance"
author: "Stace Beaulieu"
date: "2023-05-31"
output: html_notebook
---

Libraries used

```{r}
# two of the required packages are installed from GitHub
# library(remotes)
# remotes::install_github("EDIorg/EMLassemblyline")
# remotes::install_github("WHOIGit/ediutilities")

library(EMLassemblyline)
library(ediutilities)
library(here)
library(lubridate)
library(readr) 
library(dplyr)
library(tidyr)

```

Read in data tables

```{r}
# do not commit or push the data tables
staged10_wide <- read_csv("v2/input/nes-lter-zp-abundance-335um-staged10m2.csv")
staged100_wide <- read_csv("v2/input/nes-lter-zp-abundance-335um-staged100m3.csv")
unstaged10 <- read_csv("v2/input/nes-lter-zp-abundance-335um-unstaged10m2.csv")
unstaged100 <- read_csv("v2/input/nes-lter-zp-abundance-335um-unstaged100m3.csv")

# check cruise coverage
unique(staged10_wide$cruise)
unique(unstaged10$cruise)

```
Pivot the staged tables

```{r}

# Surface area haul factor
staged10_count <- staged10_wide |> pivot_longer(cols = c("adult_count", "c5_count", "c4_count", "c3_count", "c2_count", "c1_count", "cryptopia_count", "furcilia_count", "calyptopis_count", "nauplius_count", "unknown_count"), names_to = "stage", values_to = "stage_count")
staged10_haul <- staged10_wide |> pivot_longer(cols = c("adult_10m2", "c5_10m2", "c4_10m2", "c3_10m2", "c2_10m2", "c1_10m2", "cryptopia_10m2", "furcilia_10m2", "calyptopis_10m2", "nauplius_10m2", "unknown_10m2"), names_to = "stage_10m2", values_to = "stage_abundance_10m2")
staged10_count <- staged10_count |> select("cruise", "station", "cast" ,"datetime_utc", "day_night","sample_split_factor", "volume_ml","zoo_aliquot", "haul_factor", "taxa_code", "taxa_name", "conc_10m2", "zooplankton_count","totcnt", "primary_flag","secondary_flag" ,"stage","stage_count")  
staged10_haul <- staged10_haul |> select("cruise", "station", "cast" ,"datetime_utc", "day_night","sample_split_factor", "volume_ml","zoo_aliquot", "haul_factor", "taxa_code", "taxa_name", "conc_10m2", "zooplankton_count","totcnt", "primary_flag","secondary_flag" ,"stage_10m2","stage_abundance_10m2") 
staged10_haul_cols <- staged10_haul |> select("stage_10m2","stage_abundance_10m2") 
staged10 <- cbind(staged10_count, staged10_haul_cols)
staged10$stage <- gsub("_count", "", staged10$stage)
staged10 <- staged10 |>
  select(-stage_10m2, -volume_ml) |>
  rename(date_time_utc = datetime_utc) |>
  rename(aliquot_factor = zoo_aliquot) |>
  rename(haul_factor_10m2_335 = haul_factor) |>
  relocate(haul_factor_10m2_335, .before = sample_split_factor) |>
  rename(taxon_code = taxa_code) |>
  rename(taxon_name = taxa_name) |>
  rename(taxon_abundance_10m2 = conc_10m2) |>
  rename(taxon_count = zooplankton_count) |>
  rename(total_count = totcnt) |>
  relocate(total_count, .before = taxon_code) |>
  relocate(taxon_count, .after = taxon_name) |>
  relocate(primary_flag, .after = stage_abundance_10m2) |>
  relocate(secondary_flag, .after = primary_flag)

staged10 <- staged10 |>
  mutate(haul_factor_10m2_335 = round(haul_factor_10m2_335, 3)) |>
  mutate(taxon_abundance_10m2 = round(taxon_abundance_10m2, 3)) |>
  mutate(stage_abundance_10m2 = round(stage_abundance_10m2, 3))

staged10$secondary_flag <- gsub(",", "", staged10$secondary_flag)

write_csv(staged10, "nes-lter-zooplankton-transect-abundance-335um-staged10m2.csv")

# get header and stage names for metadata template
# headers_staged10 <- as.data.frame(names(staged10))
# write_csv(headers_staged10, "headers-staged10m2.csv")
# stages <- as.data.frame(unique(staged10$stage))
# write_csv(stages, "stages.csv")

```

```{r}

staged100_count <- staged100_wide |> pivot_longer(cols = c("adult_count", "c5_count", "c4_count", "c3_count", "c2_count", "c1_count", "cryptopia_count", "furcilia_count", "calyptopis_count", "nauplius_count", "unknown_count"), names_to = "stage", values_to = "stage_count")
staged100_haul <- staged100_wide |> pivot_longer(cols = c("adult_100m3", "c5_100m3", "c4_100m3", "c3_100m3", "c2_100m3", "c1_100m3", "cryptopia_100m3", "furcilia_100m3", "calyptopis_100m3", "nauplius_100m3", "unknown_100m3"), names_to = "stage_100m3", values_to = "stage_abundance_100m3")
staged100_count <- staged100_count |> select("cruise", "station", "cast" ,"datetime_utc", "day_night","sample_split_factor", "volume_ml","zoo_aliquot", "haul_factor", "taxa_code", "taxa_name", "conc_100m3", "zooplankton_count","totcnt", "primary_flag","secondary_flag" ,"stage","stage_count")  
staged100_haul <- staged100_haul |> select("cruise", "station", "cast" ,"datetime_utc", "day_night","sample_split_factor", "volume_ml","zoo_aliquot", "haul_factor", "taxa_code", "taxa_name", "conc_100m3", "zooplankton_count","totcnt", "primary_flag","secondary_flag" ,"stage_100m3","stage_abundance_100m3") 
staged100_haul_cols <- staged100_haul |> select("stage_100m3","stage_abundance_100m3") 
staged100 <- cbind(staged100_count, staged100_haul_cols)
staged100$stage <- gsub("_count", "", staged100$stage)
staged100 <- staged100 |>
  select(-stage_100m3, -volume_ml) |>
  rename(date_time_utc = datetime_utc) |>
  rename(aliquot_factor = zoo_aliquot) |>
  rename(haul_factor_100m3_335 = haul_factor) |>
  relocate(haul_factor_100m3_335, .before = sample_split_factor) |>
  rename(taxon_code = taxa_code) |>
  rename(taxon_name = taxa_name) |>
  rename(taxon_abundance_100m3 = conc_100m3) |>
  rename(taxon_count = zooplankton_count) |>
  relocate(taxon_count, .after = taxon_name) |>
  rename(total_count = totcnt) |>
  relocate(total_count, .before = taxon_code) |>
  relocate(primary_flag, .after = stage_abundance_100m3) |>
  relocate(secondary_flag, .after = primary_flag)

staged100 <- staged100 |>
  mutate(haul_factor_100m3_335 = round(haul_factor_100m3_335, 3)) |>
  mutate(taxon_abundance_100m3 = round(taxon_abundance_100m3, 3)) |>
  mutate(stage_abundance_100m3 = round(stage_abundance_100m3, 3))

staged100$secondary_flag <- gsub(",", "", staged100$secondary_flag)

write_csv(staged100, "nes-lter-zooplankton-transect-abundance-335um-staged100m3.csv")

```
Organize columns in unstaged tables

```{r}

unstaged10 <- unstaged10 |>
  select(-volume_ml) |>
  rename(date_time_utc = datetime_utc) |>
  rename(aliquot_factor = zoo_aliquot) |>
  rename(haul_factor_10m2_335 = haul_factor) |>
  relocate(haul_factor_10m2_335, .before = sample_split_factor) |>
  rename(taxon_code = taxa_code) |>
  rename(taxon_name = taxa_name) |>
  rename(taxon_abundance_10m2 = conc_10m2) |>
  rename(taxon_count = zooplankton_count) |>
  relocate(taxon_count, .after = taxon_name) |>
  rename(total_count = totcnt) |>
  relocate(total_count, .before = taxon_code) 

unstaged10 <- unstaged10 |>
  mutate(haul_factor_10m2_335 = round(haul_factor_10m2_335, 3)) |>
  mutate(taxon_abundance_10m2 = round(taxon_abundance_10m2, 3)) 

unstaged10$secondary_flag <- gsub(",", "", unstaged10$secondary_flag)

write_csv(unstaged10, "nes-lter-zooplankton-transect-abundance-335um-unstaged10m2.csv")

# get header and stage names for metadata template
# headers_unstaged10 <- as.data.frame(names(unstaged10))
# write_csv(headers_unstaged10, "headers-unstaged10m2.csv")

unstaged100 <- unstaged100 |>
  select(-volume_ml) |>
  rename(date_time_utc = datetime_utc) |>
  rename(aliquot_factor = zoo_aliquot) |>
  rename(haul_factor_100m3_335 = haul_factor) |>
  relocate(haul_factor_100m3_335, .before = sample_split_factor) |>
  rename(taxon_code = taxa_code) |>
  rename(taxon_name = taxa_name) |>
  rename(taxon_abundance_100m3 = conc_100m3) |>
  rename(taxon_count = zooplankton_count) |>
  relocate(taxon_count, .after = taxon_name) |>
  rename(total_count = totcnt) |>
  relocate(total_count, .before = taxon_code) 

unstaged100 <- unstaged100 |>
  mutate(haul_factor_100m3_335 = round(haul_factor_100m3_335, 3)) |>
  mutate(taxon_abundance_100m3 = round(taxon_abundance_100m3, 3)) 

unstaged100$secondary_flag <- gsub(",", "", unstaged100$secondary_flag)

write_csv(unstaged100, "nes-lter-zooplankton-transect-abundance-335um-unstaged100m3.csv")

```

Generate basic summary of data table

```{r}
# Just for inspecting the summary: change all character columns to factor
S10 <- staged10
S10[sapply(S10, is.character)] <- lapply(S10[sapply(S10, is.character)], as.factor)
summary(S10)

US10 <- unstaged10
US10[sapply(US10, is.character)] <- lapply(US10[sapply(US10, is.character)], as.factor)
summary(US10)
# summary(data_table)

```

Read the Excel metadata template and generate text templates used by
EMLassemblyline

```{r}

excel_template <- 'nes-lter-zooplankton-transect-abundance-info.xlsx'

sheet_to_tsv(excel_template, 'ColumnHeadersStaged10m2', 'attributes_nes-lter-zooplankton-transect-abundance-335um-staged10m2.txt') 
sheet_to_tsv(excel_template, 'ColumnHeadersStaged100m3', 'attributes_nes-lter-zooplankton-transect-abundance-335um-staged100m3.txt')
sheet_to_tsv(excel_template, 'CategoricalVariables','catvars_nes-lter-zooplankton-transect-abundance-335um-staged10m2.txt')
sheet_to_tsv(excel_template, 'CategoricalVariables', 'catvars_nes-lter-zooplankton-transect-abundance-335um-staged100m3.txt')
 
sheet_to_tsv(excel_template, 'ColumnHeadersUnstaged10m2', 'attributes_nes-lter-zooplankton-transect-abundance-335um-unstaged10m2.txt') 
sheet_to_tsv(excel_template, 'ColumnHeadersUnstaged100m3', 'attributes_nes-lter-zooplankton-transect-abundance-335um-unstaged100m3.txt')
sheet_to_tsv(excel_template, 'CategoricalVariables','catvars_nes-lter-zooplankton-transect-abundance-335um-unstaged10m2.txt')
sheet_to_tsv(excel_template, 'CategoricalVariables', 'catvars_nes-lter-zooplankton-transect-abundance-335um-unstaged100m3.txt')

sheet_to_tsv(excel_template, 'Personnel', 'personnel.txt')

sheet_to_tsv(excel_template, 'Keywords', 'keywords.txt')

sheet_to_tsv(excel_template, 'CustomUnits', 'custom_units.txt')

```

Use templates in "template core metadata" step from EML assembly line

```{r}

template_core_metadata(path=here(), license='CCBY', file.type = ".md")

datatables <- c('nes-lter-zooplankton-transect-abundance-335um-staged10m2.csv', 'nes-lter-zooplankton-transect-abundance-335um-staged100m3.csv', 'nes-lter-zooplankton-transect-abundance-335um-unstaged10m2.csv', 'nes-lter-zooplankton-transect-abundance-335um-unstaged100m3.csv' )
descriptions <- c('Staged zooplankton count and abundance at 10m2 haul factor', 'Staged zooplankton count and abundance at 100m3 haul factor', 'Unstaged zooplankton count and abundance at 10m2 haul factor', 'Unstaged zooplankton count and abundance at 100m3 haul factor')
tablenames <- c('abundance-335um-staged10m2', 'abundance-335um-staged100m3', 'abundance-335um-unstaged10m2', 'abundance-335um-unstaged100m3')

```
Generate the package and insert the parent project node into the resulting EML

```{r}
# package ID 25
pkg_id <- 'knb-lter-nes.25.2'

make_eml(here(),
         dataset.title='Zooplankton abundance from net tows on Northeast U.S. Shelf Long Term Ecological Research (NES-LTER) Transect cruises, ongoing since 2018',
        data.table= datatables,  
         data.table.description = descriptions,
         data.table.name = tablenames,
         other.entity = 'nes-lter-zp-cruises-metadata.csv',
         other.entity.name = 'List of NES-LTER cruises',
         other.entity.description = 'List of NES-LTER cruises',         
         temporal.coverage = temporal_coverage(staged10$date_time_utc),
         #temporal.coverage = c('2018-01-31', '2020-10-18'), # from NES-LTER_TOW_data
         geographic.description = "NES-LTER Transect",
         #         geographic.coordinates = geographic_coordinates(data_table$latitude, data_table$longitude),
         geographic.coordinates = c(41.5240, -70.8739, 39.7596, -71.1344), # from NES-LTER_TOW_data in v1
         maintenance.description = "ongoing",
         user.id = "NES",
         user.domain = "LTER",
         package.id = pkg_id)

project_insert(pkg_id,'parent_project_NESI-II_RAPID.xml')
```
